Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-23 Manuscript/manuscript.Rmd
manuscript edits

## LOW IMBALANCE RATIOS ACROSS RESOLUTIONS & CELL LINES

After classifying each genomic bin as either being flanked by a TAD boundary or not, it was found that the subsequent imbalance ratios across resolutions and cell lines were extremely low. Ideally a manageable dataset will have an imbalance ratio hovering near 1. At their heights, the imbalance ratios were 0.179, 0.152, ... for cell lines GM12878, K562, ... respectively (Figure 6). The imbalance ratio were lowest for the 10 kb resolution data for both GM12878 and K562.

@@

We first evaluated which predictor types were best able to discriminate between genomic bins that either flanked or did not flank TAD boundaries. To do so we performed a series of random forest classification algorithms on the full features space (all combined predictor types), as well as on type-specific feature spaces (overlap counts, overlap predictors, and log2 distances). It was found that models with distance type predictors had higher AUPRC values compared to other predictor types (Figure 4). This result was observed across both data resolution and cell line, with the largest AUPRC obtained from the 10 kb resolutions for all [gm12878, k562, ???update as more cell lines included] cell lines (0.431 and 0.456 respectively). There was a gradual decline in AUPRC values as resolution increased with models built on 100 kb resolution domain data having the lowest AUPRC curves, regardless of predictor type (Supplementary Table 3).

Recursive feature elimination was used to establish the optimal number of distance-type features to include in downstream analyses. It was determined that performances according to AUPRC values were asymptotically similiar for random forest models with sets of predictors greater than 32 (Figure 5). Therefore, we reduced the feature space for subsequent models to the top 32 predictors, respective to resolution and cell line. The sets of top predictors chosen varied across resolution and cell line due to differences in the parameters provided in the TAD calling algorithm.

>>>>

@@

Prior to resampling, we performed a series of random forest classification algorithms on the full features space (all combined predictor types), as well as on type-specific feature spaces (overlap counts, overlap predictors, and log2 distances) to evaluate which predictor types were best able to discriminate between genomic bins that either flanked or did not flank TAD boundaries. It was found that, for 10 kb resolution data, models with distance type predictors had higher AUPRC values compared to other predictor types for both GM12878 and K562 cell lines (Figure 4). This result was observed across all other data resolutions as well (Supplementary Figure 2). There was a gradual decline in AUPRC values as resolution increased with models built on 100 kb resolution domain data having the lowest AUPRC curves, regardless of predictor type (Supplementary Table 3).

Recursive feature elimination was used to establish the optimal number of distance-type features to include in downstream analyses. It was determined that, for 10 kb resolution data, near optimal performance was obtained for models with only 32 predictors for each cell line (Figure 5). Results remained consistent across the other data resolutions (Supplementary Figure 4). Therefore, we reduced the feature space for subsequent models to the top 32 predictors, respective to resolution and cell line. The sets of top predictors chosen varied across resolution and cell line due to differences in the parameters provided in the TAD calling algorithm. For the 10 kb resolution data, the most important predictive features that were robust across cell lines were

## DATA RESAMPLING IMPROVES MODEL PREDICTION

The low baseline thresholds represented by the dashed line in the precision-recall curves (Figure 4 and Supplementary Figure2) indicate the wide disparity between majority and minority classes. Ideally a manageable dataset will have an imbalance ratio hovering near 1. However, in our study we found that there was an average imbalance ratio of 0.154 (SD=0.022) and 0.137 (SD=0.0181) for cell lines GM12878 and K562 respectively (Supplementary Figure 3).  This highlights the need for suitable resampling techniques.
